---
title: "Module-4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Module 4

These are my records and notes of Module 4.

### Challenge 1:
First, create a vector of 1 word character strings comprising the first line of the Ulysses, by James Joyce: “Stately, plump Buck Mulligan came from the stairhead, bearing a bowl of lather on which a mirror and a razor lay crossed.”


1. rename sentence 
2. remove punctuation
3. Split sting, "strsplit"

```{r split}
x<-"Stately, plump Buck Mulligan came from the stairhead, bearing a bowl of lather on which a mirror and a razor lay crossed." #must use R quotes not sentence quotes
x2<-((gsub("[[:punct:]]","",x))) #does not remove quote marks
xspt<-c(unlist(strsplit(x2," "))) #creates as a list, so used unlist function
xspt
```

Then, extract every third element of this vector.
```{r extract}
extract<- xspt[seq(from=3, to=22, by = 3)]
extract
```

HOORAY!

###Challenge 2

```{r matrix, echo=TRUE}
m <- matrix(data = 1:80, nrow = 8, ncol = 10, byrow = FALSE)
m
```
1. Given the matrix, m, above, extract the 2nd, 3rd, and 6th columns and assign them to the variable x
```{r extract columns}
x<-m[,c(2,3,6)]
x
```
2.Given the matrix, m, above, extract the 6th to 8th row and assign them to the variable x
```{r extract rows}
x<-m[c(6:8),]
x
```

3. Given the matrix, m, above, extract the elements from row 2, column 2 to row 6, column 9 and assign them to the variable x
```{r extract matrix}
x<-m[c(2:6),c(2:9)]
x
```

### Challenge 3

Construct a 4-dimensional, 400 element array (5 x 5 x 4 x 4) named a consisting of the numbers 400 to 1 (i.e., a descending series)
```{r array}
a<- array(data=400:1, dim=c(5,5,4,4))
a
```

Given this matrix, what would the following return?

```{r array-call}
a[1, 1, 1, 2]

a[2, 3, 2, ]

a[1:5, 1:5, 3, 3]

```
### challenge 4
Create a list representing this (simplified) primate taxonomy. HINT: you can use lists as elements in a list.

```{r list}
Plat<-c("Cebidae", "Atelidae", "Pitheciidae")
Cat<- c("Cercopithecidae","Hylobatidae","Hominidae")
Tar<-c("Tarsiidae")
Lor<-c("Lorisidae", "Galagidae")
Lem<-c("Cheirogaleidae", "Lepilemuridae", "Indriidae", "Lemuridae", "Daubentoniidae")
Anth<-list(Plat, Cat)
names(Anth)<-c("Platyrrhini","Catarrhini")
Hap<-list(Anth, Tar)
names(Hap)<-c("Anthropoidea", "Tarsioidea")
Strep<-list(Lor, Lem)
names(Strep)<-c("Lorisoidea","Lemuroidea")
primates<-list(Hap, Strep)
names(primates)<-c("Haplorhini", "Strepsirhini")
primates
```


###challenge 5
Store the following numbers as a 5 x 3 matrix: 3, 0, 1 ,23, 1, 2, 33, 1, 1, 42, 0, 1, 41, 0, 2. Be sure to fill the matrix ROWWISE.
```{r matrix2}
d<- c(3, 0, 1 ,23, 1, 2, 33, 1, 1, 42, 0, 1, 41, 0, 2)
mat<- matrix(data=d, 5, 3, byrow=TRUE)
mat
```
Coerce the matrix to a data frame.
```{r}
df<-data.frame(mat)
df
```
As a data frame, coerce the second column to be logical-valued
```{r}
df[,2] <-as.logical (df[,2])
df
```
As a data frame, coerce the third column to be factor-valued
```{r}
df[,3]<-as.factor(df[,3])
```

When you are done, use the str() command to show the data type for each variable in your dataframe.
```{r}
str(df)
```
